---
title: "Premier League Stats"
description: |
  A look at the 22/23 Premier League season so far.
author:
  - name: Mark Johnson
    url: 
date: 2023-01-12
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(gt)
```


> 2022/2023 English Premier League Season 

So far the English Premier League season is approximately half way through. Lets create a table for each premier league team displaying the highest number of shots, the game week that occurred and their opponent.

I'll be using free Premier League data found here - https://www.football-data.co.uk/englandm.php

```{r}

epl <- read.csv("C:/Users/15593/Documents/GitHub/Mark-s-blog/E0.csv") %>% 
  mutate(key = paste(AwayTeam, HomeTeam, Date))

epl_away <- epl %>% 
  select(AwayTeam, AS, Date, key, HomeTeam) %>% 
  rename(Team=AwayTeam, Shots=AS) %>% 
  mutate(status="away") %>% 
  rename(Oponent=HomeTeam)

epl_home <- epl %>% 
  select(HomeTeam, HS, Date, key, AwayTeam) %>% 
  rename(Team=HomeTeam, Shots=HS) %>% 
  mutate(status="home") %>% 
  rename(Oponent=AwayTeam)

epl_both <- bind_rows(epl_home, epl_away)

epl_max_shots <- epl_both %>% 
  group_by(Team) %>% 
  top_n(1, Shots) %>% 
  arrange(Team) %>% 
  select(-key)



epl_table <- gt(epl_max_shots)

epl_table
  

```

Now lets see which team has given up the most shots so far

```{r, echo=FALSE}

worst_shots <- epl_max_shots %>% 
  group_by(Oponent) %>% 
  count() %>% 
  arrange(Oponent) %>% 
  rename(Team=Oponent) %>% 
  arrange(desc(n))

gt(worst_shots)

```


From the table we can see Everton has given up the most shots to five other teams so far in the 22/23 season. 




Hi Lauren!